Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE1)

### 2014-04-24 WSSSPE1_report/WSSSPE1_summary_report.tex
comment a note
minor fix in abstract

\begin{abstract} %\note{Lead: Sou-Cheng T. Choi}
This document summarizes the First Workshop on Sustainable Software
for Science: Practice and Experiences (WSSSPE1). The content is based
on a set of collaborative notes taken during the workshop, including
discussion about the keynote presentations, three themed panels that
discussed papers accepted at the workshop`, other discussion,
cross-cutting issues, use cases, and conclusions. It joins a nascent
literature seeking to understand what drives software work in science
and how the reward systems of science function, thereby shaping the
types of software projects undertaken, including the extent to which
developers are motivated to build software for the long-term, for the
use of others, and whether to work collaboratively or separately.
Unique perspectives were captured about issues such as comprehensive
documentation, development and deployment practices; software
licenses; career paths for developers; identifiers for software
through a mechanism such as a Digital Object Identifier or publication
of a ``software paper''; and evidence of software contribution and
impact, recorded in online systems such as source code repositories
like GitHub. For truly sustainable software, there should be no
endpoint, as the software products continue to be used beyond a single
institution, grant, and developer or development team.

@@

% Why are we classifying software like above?
This classification is motivated by our argument that the two kinds of
software projects wildly vary in factors such as scope, purpose and
usage. The development and management of each kind is significantly
different. Consequently, the sustainability challenges faced by them
differ and must be treated separately. For instance, the challenges
faced by a gateway software development project such as
\toolname{CIPRES} or visualization software products such as
\toolname{VisIT} or \toolname{ParaView} are distinct to a niche software for
\textit{ab initio} modeling and simulation such as \toolname{VASP} or
\toolname{Quantum Espresso}. \note{cite SW}

>>>>

\begin{abstract} %\note{Lead: Sou-Cheng T. Choi}
This document summarizes the First Workshop on Sustainable Software
for Science: Practice and Experiences (WSSSPE1). The content is based
on a set of collaborative notes taken during the workshop, including
discussion about the keynote presentations, three themed panels that
discussed papers accepted at the workshop, other discussion,
cross-cutting issues, use cases, and conclusions. It joins a nascent
literature seeking to understand what drives software work in science
and how the reward systems of science function, thereby shaping the
types of software projects undertaken, including the extent to which
developers are motivated to build software for the long-term, for the
use of others, and whether to work collaboratively or separately.
Unique perspectives were captured about issues such as comprehensive
documentation, development and deployment practices; software
licenses; career paths for developers; identifiers for software
through a mechanism such as a Digital Object Identifier or publication
of a ``software paper''; and evidence of software contribution and
impact, recorded in online systems such as source code repositories
like GitHub. For truly sustainable software, there should be no
endpoint, as the software products continue to be used beyond a single
institution, grant, and developer or development team.

@@

% Why are we classifying software like above?
This classification is motivated by our argument that the two kinds of
software projects wildly vary in factors such as scope, purpose and
usage. The development and management of each kind is significantly
different. Consequently, the sustainability challenges faced by them
differ and must be treated separately. For instance, the challenges
faced by a gateway software development project such as
\toolname{CIPRES} or visualization software products such as
\toolname{VisIT} or \toolname{ParaView} are distinct to a niche software for
\textit{ab initio} modeling and simulation such as \toolname{VASP} or
\toolname{Quantum Espresso}. %\note{cite SW}
